### 2024-11-04 paper/paper.qmd
Update paper.qmd
Added survey questions and refined appendix.

---
title: "Forecasting the 2024 U.S. Presidential Election"
subtitle: "Using Aggregated Polling Data to Predict the Outcome of the 2024 Presidential Race"
author: 
  - Andy Jiang
  - Onon Burentuvshin
  - Rui Hu
thanks: "Code and data are available at: [https://github.com/AndyYanxunJiang/president-polls)."
date: today
date-format: long
abstract: "The 2024 U.S. presidential election is shaping up to be a highly swing race, with polling data playing a crucial role in gauging public sentiment and forecasting the potential outcome. This paper aims to build a predictive model using a poll-of-polls approach, aggregating data from multiple pollsters to forecast the winner of the upcoming election. The analysis incorporates polling data from various sources, filtered to include only high-quality pollsters with a numeric grade above 3, to ensure the reliability of the predictions. Additionally, a detailed examination of one selected pollster’s methodology is conducted, highlighting strengths, weaknesses, and biases in the survey process. The paper concludes with an idealized polling methodology, designed within a $100K budget, to enhance the accuracy of election forecasts. Results are presented in the context of understanding polling trends, potential biases, and the evolving landscape of voter preferences as the election approaches."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
library(tidyverse)
library(readr)
library(here)
library(knitr)
library(kableExtra)
library(scales)
library(arrow)
library(bayesplot)
library(rstanarm)

data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
```

# Introduction
The 2024 U.S. presidential election is shaping up to be one of the most closely watched and potentially pivotal elections in recent history. As public opinion shifts and political dynamics evolve, polling data has emerged as a critical tool for forecasting the outcome and understanding voter preferences. Polls provide a snapshot of where candidates stand at various points in time, reflecting the impact of campaigns, debates, and major events. However, the reliability and accuracy of individual polls can vary significantly based on methodology, sample size, and other factors, making it necessary to aggregate multiple sources of data for a more robust prediction.

This paper aims to forecast the outcome of the 2024 U.S. presidential election by leveraging a "poll-of-polls" approach, which aggregates data from various pollsters to create a more comprehensive picture of the race. By focusing on high-quality polls and conducting a detailed analysis of the polling trends for the two main candidates, Donald Trump and Kamala Harris, we seek to uncover patterns in voter support and potential factors influencing shifts in public opinion. Additionally, the paper examines the methodologies of different pollsters to assess their strengths and weaknesses, which can help identify possible biases in the aggregated data.

The analysis will further explore the geographic distribution of support across key swing states, providing insights into the electoral landscape. Finally, the paper presents an idealized polling methodology with a $100,000 budget, designed to improve the accuracy and representativeness of election forecasts. By integrating multiple data sources and critically evaluating polling practices, this study aims to contribute to the ongoing discussion on the role of polling in democratic processes and election forecasting.

# Data {#sec-data}
The data used in this paper was obtained from the FiveThirtyEight website [@Best2024], which aggregates polling data and other relevant indicators to track the U.S. election landscape. We accessed this data using custom web scraping tools and APIs provided by FiveThirtyEight. The dataset includes information on pollster, state, start and end dates of polls, sample size, population type (e.g., registered voters or likely voters), and party affiliation.

## Overview
Our analysis was conducted using R [@citeR] along with several supporting packages. We used the **tidyverse** for data manipulation [@tidyverse], **readr** for data import [@readr], and **here** for file management [@here].

Dynamic reporting was handled with **knitr** [@knitr], while **kableExtra** allowed for enhanced table formatting [@kableExtra]. We used **scales** to adjust visualization scales [@scales], and **arrow** for efficient data processing [@arrow].

For Bayesian modeling and visualization, we used **bayesplot** [@bayesplot] and **rstanarm** [@rstanarm].

## Measurement
The FiveThirtyEight dataset is updated in real-time, providing the latest polling data for the 2024 U.S. presidential election. This dataset aggregates information from multiple reputable polling sources, applying adjustments to enhance reliability, such as weighting based on pollster quality and recency of data. These measures ensure that the dataset remains a dependable reflection of current voter sentiment and trends as Election Day approaches.

## Analysis Data
The analysis dataset includes several key variables: State, Sample Size, Population Type (e.g., Registered Voters), Party Affiliation, Candidate Preference (Trump or Harris), and Pollster. These variables allow for an analysis of voter sentiment in specific swing states, with a focus on support levels between Donald Trump and Kamala Harris.

Certain variables, such as Question ID, Start Date, End Date, and Hypothetical Questions, were removed as they were not necessary for this analysis. Additionally, data on other candidates was excluded to concentrate solely on the Trump vs. Harris matchup, aligning with the focus on the most relevant contenders in the seven swing states: Pennsylvania, Georgia, North Carolina, Michigan, Arizona, Wisconsin, Nevada.

@tbl-sample provides a random sample of the analysis dataset, displaying all relevant variables used for analysis.

```{r}
#| label: tbl-sample
#| tbl-cap: Random sample of 10 rows from the analysis data.
#| echo: false
#| warning: false
#| message: false

set.seed(311)

# Sample 10 random observations from the filtered data with Trump and Harris, and remove unnecessary columns
sample_data <- data %>%
  filter(candidate_name %in% c("Donald Trump", "Kamala Harris")) %>%
  sample_n(10) %>%
  select(-start_date, -end_date, -question_id, -hypothetical, -pct)

# Display the sampled data as a table with refined styling
kable(sample_data, caption = "Sample of 10 Random Observations from the Analysis Data") %>%
  kable_styling(full_width = FALSE, position = "center") %>%  # Center the table in the PDF
  column_spec(1, width = "3cm") %>%    # Adjust specific columns for better fit
  column_spec(2, width = "2.5cm") %>%
  column_spec(3, extra_css = "padding-left: -5px;") %>%  # Move sample_size column slightly to the left
  column_spec(4:5, width = "1.5cm")    # Adjust remaining columns with less text

```

## Predictor variables

# Model
In this analysis, we utilize a logistic regression model to predict polling percentages for the 2024 U.S. presidential candidates over time. In order to focus on states with historical significance in elections. The model is specified as a Generalized Linear Model (GLM) with a binomial family and a logit link function, appropriate for modeling binary or proportion-based outcomes.

## Model set-up
The dataset, model_data.csv, provides polling data with average_pct as the observed average polling percentage among 7 swing states. The end_date indicates the polling period's endpoint.
To predict polling percentages beyond the available data, we created an additional sequence of 20 days beyond the dataset's last recorded date. Because we want to predict the result in November 5. This expansion allows for forecasting near-term polling trends. Also, capturing potential shifts in candidate support leading up to election day. The model was trained on existing data and applied to this extended date range using predict(model, newdata = ., type = "response").

Our model setup and variable selection are as follows:
$$
\text{logit}(p_i) = \beta_0 + \beta_1 (\text{Date}) + \beta_2 (\text{Candidate})
$$

where:

- $$p_i \ \text{represents the probability of support for candidate i at a specific date.}$$
- $$\beta_0 \ \text{is the intercept, representing baseline support.}$$
- $$\beta_1 \ \text{captures changes in support over time (date effect).}$$
- $$\beta_2 \ \text{is a categorical variable accounting for each candidate.}$$

### Model justification
The choice of a logistic regression model was guided by the nature of polling data. Nature of polling data represents proportions that naturally bound between 0 and 1. The logistic function’s S-shaped curve effectively models the relationship between time and polling percentage, accommodating non-linear trends that may arise as election day nears.
Firstly, candidate names should be included as a categorical predictor in the model. It allows us to capture individual impact in support rates by candidate. In addition, a transformation from date to numeric is made on end_date in order to get a numerical format. It is called the end_date_numeric. It helps put date as a predictor in the model. By structuring the model this way, it can predict the percentage of the poll by both the impact of candidates and trends over time.

# Results
The analysis of polling data across key swing states for the 2024 U.S. Presidential Election reveals insights into the levels of support for candidates Donald Trump and Kamala Harris. The findings are presented through three primary visualizations: the average polling percentage by state, the distribution of polling percentages, and the difference in average polling percentage between the two candidates.

## Average Polling Percentage by State
From @fig-1, we can see that in the swing states of Arizona, Georgia, and North Carolina, Trump holds a narrow lead. In contrast, Harris leads in Pennsylvania, Wisconsin, Nevada, and Michigan. When looking at the overall average across these states, Harris slightly edges out Trump, indicating her overall advantage in these swing areas.

## Polling Percentage Distribution
The box plot in @fig-2 demonstrates the range and consistency of support for both candidates across each state. Trump shows a wider range of support, reflecting variability, while Harris has relatively steady polling percentages across states, maintaining higher consistency. Harris’s support is less variable and generally centers around higher polling percentages compared to Trump’s.

## Difference in Average Polling Percentage
@fig-3 highlights the difference in average polling percentages by state, with positive values indicating Trump’s lead and negative values indicating Harris's. Trump’s lead in the three states (Arizona, Georgia, and North Carolina) is relatively narrow, with margins of less than 1 percent. In contrast, Harris’s lead in the states favoring her (Pennsylvania, Wisconsin, Nevada, and Michigan) ranges from 1 to 4 percent, indicating her stronger support in the states where she leads.

```{r}
#| label: fig-1
#| fig-cap: "Average Polling Percentage by State"
#| echo: false
#| warning: false
#| message: false
# Calculate the average polling percentage by state for Trump and Harris
data_avg <- data %>%
  filter(candidate_name %in% c("Donald Trump", "Kamala Harris")) %>%
  group_by(state, candidate_name) %>%
  summarize(avg_pct = mean(pct, na.rm = TRUE), .groups = "drop")

# Calculate the overall average polling percentage for Trump and Harris
overall_avg <- data %>%
  filter(candidate_name %in% c("Donald Trump", "Kamala Harris")) %>%
  group_by(candidate_name) %>%
  summarize(avg_pct = mean(pct, na.rm = TRUE)) %>%
  mutate(state = " Overall")  # Add a space before "Overall" to separate it visually

# Combine state-level data with overall average
combined_data <- bind_rows(data_avg, overall_avg)

# Ensure "Overall" appears at the bottom with spacing
combined_data$state <- factor(combined_data$state, levels = c(setdiff(unique(combined_data$state), " Overall"), " Overall"))

# Plot the average polling percentages by state with overall comparison
ggplot(combined_data, aes(x = state, y = avg_pct, fill = candidate_name)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), aes(color = ifelse(state == " Overall", "Overall", "State")), width = 0.6) + 
  scale_color_manual(values = c("Overall" = "black", "State" = NA), guide = "none") +  # Add dark grey outline for "Overall" bars
  scale_fill_manual(values = c("Donald Trump" = "#F8766D", "Kamala Harris" = "#56B4E9")) +  # Custom colors for consistency
  labs(title = "Average Polling Percentages by State",
       x = "State",
       y = "Average Polling Percentage",
       fill = "Candidate") +
  coord_flip() +  # Flip coordinates for easier reading
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.y = element_text(margin = margin(r = 10)), # Add spacing to y-axis labels
    panel.grid.major.y = element_blank(), # Remove y-axis gridlines for clarity
    panel.spacing = unit(1, "lines") # Increase panel spacing
  )

```

```{r}
#| label: fig-2
#| fig-cap: "Polling Percentage Distribution"
#| echo: false
#| warning: false
#| message: false
ggplot(data %>% filter(candidate_name %in% c("Donald Trump", "Kamala Harris")), 
       aes(x = state, y = pct, fill = candidate_name)) +
  geom_boxplot(alpha = 0.7) +
  labs(title = "Polling Percentage Distribution by State",
       x = "State",
       y = "Polling Percentage",
       fill = "Candidate") +
  scale_fill_manual(values = c("Donald Trump" = "#F8766D", "Kamala Harris" = "#56B4E9")) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
```{r}
#| label: fig-3
#| fig-cap: "Difference in Average Polling Percentage"
#| echo: false
#| warning: false
#| message: false
data_diff <- data %>%
  filter(candidate_name %in% c("Donald Trump", "Kamala Harris")) %>%
  group_by(state) %>%
  summarize(diff_pct = mean(pct[candidate_name == "Donald Trump"], na.rm = TRUE) - 
                      mean(pct[candidate_name == "Kamala Harris"], na.rm = TRUE), .groups = "drop")

ggplot(data_diff, aes(x = reorder(state, diff_pct), y = diff_pct, fill = diff_pct > 0)) +
  geom_bar(stat = "identity", color = "darkgrey") +
  scale_fill_manual(values = c("TRUE" = "#F8766D", "FALSE" = "#56B4E9"), 
                    labels = c("Kamala Harris", "Donald Trump")) +
  labs(title = "Difference in Average Polling Percentage by State",
       x = "State",
       y = "Difference in Polling Percentage",
       fill = "Lead") +
  coord_flip() +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Swing State Polling Prediction Model
@fig-4 shows the trend of support rate changes for each candidate before the upcoming election, as well as the model's predicted values for the next 20 days.
```{r}
#| label: fig-4
#| fig-cap: "Swing State Polling Prediction Model"
#| echo: false
#| warning: false
#| message: false

invisible(Sys.setlocale("LC_TIME", "en_US"))
data_filtered <- read_parquet(here::here("data/02-analysis_data/model_data.parquet"))
model <- readRDS(here::here("models/model.rds"))
data_filtered <- data_filtered %>%
  mutate(predicted_pct = predict(model, newdata = ., type = "response"))

last_date <- max(data_filtered$end_date, na.rm = TRUE)
new_dates <- seq.Date(from = last_date + 1, by = "day", length.out = 20)
new_data <- expand.grid(
  end_date = new_dates,
  candidate_name = unique(data_filtered$candidate_name)
)
new_data <- new_data %>%
  mutate(
    end_date_numeric = as.numeric(end_date), 
  ) %>% mutate(predicted_pct = predict(model, newdata = ., type = "response"))
all_data <- bind_rows(data_filtered, new_data)

ggplot(all_data, aes(x = end_date, y = pct, color = candidate_name)) +
  geom_point(aes(y = average_pct/100), alpha = 0.5) + 
  geom_line(aes(y = predicted_pct), size = 1) + 
  labs(title = "2024 US President Polls - Predicted vs True",
       x = "Date",
       y = "Rate of support") +
  theme_minimal() +
  scale_x_date(date_labels = "%b", date_breaks = "1 month") +
  scale_y_continuous(labels = percent_format(), limits = c(0.44, 0.48))

```

# Discussion

## Impact of Key Demographics and Policy Appeals
Harris’s lead in several swing states may be influenced by her campaign’s appeal to key demographic groups, particularly immigrants. The Democratic platform has emphasized social and economic benefits for immigrant communities, which may have strengthened her base in states with significant immigrant populations. Policies that promise support for healthcare, education, and other public services are likely contributing to Harris’s appeal among these groups, solidifying her lead in states like Nevada and Michigan. This demographic support provides Harris with a strategic advantage, as these communities may play a crucial role in voter turnout and loyalty.

## Influence of Endorsements and High-Profile Backing
Recently, high-profile figures like Elon Musk have publicly thrown their support behind Trump, leveraging their influence to rally support for his campaign [@Zahn2024]. Musk’s endorsement could invigorate Trump’s base, especially among Musk’s followers and those engaged in the fields of business and technology. This visible backing from such influential figures has the potential to sway undecided voters and influence last-minute decisions, particularly in swing states where Trump’s lead remains narrow.

Musk’s endorsement might not only energize Trump’s existing supporters but could also lead to a boost in campaign funds and social media engagement, further strengthening Trump’s position. In closely contested states, where a few thousand votes can make a significant difference, this level of backing from a high-profile figure could be crucial. If Musk’s involvement draws in additional resources and amplifies Trump’s visibility, it may ultimately impact the final outcome of the election on November 5th.

## Limitations of the Analysis: Focus on Swing States
While the analysis focuses on key swing states, it does not consider the broader national picture, which includes states with clear, one-sided advantages for either candidate. This narrowed focus on swing states provides insights into swing areas but does not give a complete picture of the national landscape. Therefore, while Harris appears to have an edge in these swing states, it is unclear whether she is leading overall. The situation in solidly red or blue states could still shift the electoral balance, meaning Trump could potentially gain ground through strong performances in traditionally Republican states.

## Potential Influence of Late Campaign Efforts
With a few days remaining before the election, both campaigns are likely to intensify their outreach efforts, particularly in swing states. Trump’s recent increase in targeted advertisements and campaign rallies in swing states could energize his supporters, narrowing Harris’s lead. On the other hand, Harris’s campaign may focus on mobilizing key demographics, such as young and minority voters, to secure her advantage. Late-breaking events, public debates, or controversies could sway undecided voters and ultimately impact the election results in these states.

## Limitations and Next Steps
This analysis is limited to examining current polling data in swing states and does not account for unexpected shifts in voter sentiment or turnout on election day. Additionally, focusing exclusively on swing states omits the influence of non-swing states that may still play a role in the electoral vote tally. Future analyses could incorporate national polling data to provide a more comprehensive perspective and examine demographic breakdowns within each state to understand which groups are driving support for each candidate. Integrating data on recent campaign activities, endorsements, and voter sentiment on pressing issues could also improve the predictive accuracy of election outcomes.

\newpage

\appendix

# Appendix A {-}

#  Sampling Methodology Employed by New York Times and Siena College:

Our analysis and prediction utilizes data collected by a single pollster and it is important to scrutinize the sampling methodology behind the sampling methods used to determine the extent of the external validity of our predictions. Primarily, the data gathered by New York Times was done so with the collaboration with Sienna College through live-interviewers reaching voters by telephone.

## Target Population and Sampling Frame
The poll targets registered voters in the United States, with a particular focus on voters in swing states for presidential races. Consequently, the target population is registered voters of the United States of America and the sampling frame is the voter file, a national list of registered voters from which NYC employs random sampling to create their pollsters. This voter file contains information on more than 200 million Americans that includes information about their party affiliation, voting history, and  basic demographic information. From this sampling frame, respondents are randomly selected while balancing key demographic and political characteristics through both front-end sampling and post-survey weighting pursuant to appropriate representation of the entire target population.

## Recruitment Methods
The poll relies on random digit dialing to reach voters, predominantly via cell phones. While landlines are also called, the dominance of cellphone usage aligns with contemporary communication trends. This results in their rendition of random sampling, which contributes to the external validity of data. However, reaching voters by phone has become more difficult due to low response rates requiring many attempts to secure responses from certain individuals (response rates are as  low as 2% at NYC). In fact, according to an article by Emerson College, telephone polling has become entirely obsolete (@EmersonCollege). Despite this, the phone methodology is still seen as effective for its ability to quickly reach a diverse voter pool across various states as well as seen as the golden standard for polling. Although this can be seen as random sampling, the concern of low response rates and reliance on dialing to reach voters raises concerns of selection bias. There are questions as to whether a certain demographic will be more responsive to being reached through a cell phone. Moreover, voters that are reached over the phone may feel the urge to provide more ‘socially desirable’ responses that are not reflective of their true opinions.

## Non-Response Remedies
With a response rate of 2%, there is valid concern over non-response bias underrepresenting certain demographics. The poll addresses non-response through a combination of front-end sampling adjustments (calling harder-to-reach voters) and post-survey weighting. Weighting is applied to ensure that the final sample aligns with the demographic and political composition of the electorate, based on voter file data. Non-response bias is a persistent challenge, and even though weighting helps, it's impossible to be certain that respondents are fully representative of those who do not respond. Experiments with alternative methods, such as mail surveys with incentives of cash vouchers have shown to increase response rate up to 30% (New York Times).

# Sampling Approach: Stratified Sampling

##  Methodology
The Times/Siena Poll employs a stratified random sampling approach to ensure that the sample accurately represents the diversity within the target population of registered voters in the United States. This method involves dividing the population into distinct subgroups, or “strata,” based on key characteristics such as state, age, gender, race, education level, and political affiliation. For example, in presidential elections, particular emphasis is placed on voters from swing states like Arizona, North Carolina, Georgia, and Pennsylvania. By focusing on these battleground states, the poll aims to capture the opinions of voters who are most likely to influence the election outcome. Once the strata are established, respondents are randomly selected within each group, ensuring that all demographic groups are appropriately represented in proportion to their actual distribution within the electorate. Furthermore, post-survey weighting is employed to more closely resemble this distribution.

## Advantages
Stratified sampling offers the advantage of higher precision in representing the population's diversity, as each subgroup is proportionally represented. By breaking down the sample into distinct groups, the Times/Siena Poll can adjust for demographic variations between states or voter groups, which is particularly useful in a diverse electorate. This approach helps to minimize sampling error by ensuring that all significant voter subgroups are included, thus increasing the reliability of the poll’s predictions. It is standard practice for them to sample 1,000 respondents for their polls. Using stratified sampling to have this sample size closely resemble that of the entire target population cuts costs and provides efficient real-time data.

## Stratified Sampling over Random Sampling

If the methodology randomly sampled from the target population without regard for the distribution of demographics in the state, significant increases to the sample sizes have to be made. Random sampling will have largely less precision in representing the distribution of different demographics within a given state unless a large enough sample size were to be drawn. Although it is ideal to have probability based sampling increasingly resemble the demographic distribution quota with growing sample sizes, it is unfeasible to conduct such a large scale poll as this approach would encounter a limitation of resources. With declining response rates and the difficulty of reaching voters, New York Times/Siena's nonprobability sampling is a more realistic and efficient method of polling than probability sampling is (@tellingstorieswithdata). As such, the stratified sampling approach allows New York Times/Siena to represent the entire target population with a sample size of 1,000.

## Challenges and Trade-offs

Although stratified sampling improves representativeness, it presents certain challenges in execution. First, identifying and reaching a sufficient number of respondents from harder-to-reach groups (such as young voters, minority populations, and those without reliable phone access) can be difficult, especially given the 2% response rate. To counterbalance the low response rates, the poll requires multiple call attempts and careful post-survey weighting to adjust for any potential underrepresentation. Additionally, while phone polling offers broad geographic reach and a standardized mode of data collection, it limits the sample to voters with phone access who are willing to participate in phone interviews, potentially introducing selection bias if certain demographics are less reachable by phone. Although New York Times/Sienna College employs random digit dialing from their sampling frame, their rendition of sample recruitment requires them to sample individuals according to quotas, resulting in nonprobability sampling.

## Evaluation of Questionnaire Design in the Times/Siena Poll

### Strengths

The Times/Siena Poll’s approach to questionnaire design reflects a careful effort to craft unbiased, relevant, and inclusive questions. By tailoring questions to current events and topics of widespread interest, the poll aims to capture public sentiment on issues that directly affect voters' attitudes, such as economic conditions or recent political debates. This responsiveness to current events strengthens the poll's relevance and allows it to track shifts in opinion over time. The poll’s commitment to neutrality is a significant strength. Each question is designed to ensure that respondents, regardless of their political views, feel that their perspectives are fairly represented. The team behind the poll spends substantial time crafting response options to reflect a balanced spectrum of views, minimizing potential response bias and enhancing the questionnaire's reliability across different demographic and ideological groups (@NewYorkTimes). Clarity is also a central focus in their questionnaire development. The team emphasizes ensuring that each question is clear and unambiguous, minimizing the risk of misinterpretation. This dedication to straightforward, widely understandable questions improves the validity of responses, as it reduces the chance that respondents interpret questions differently based on background or knowledge level.

### Limitations

Despite the strengths of the Times/Siena Poll’s questionnaire design, there are some limitations. One notable concern is that, although the poll tries to avoid leading questions, any standardized questionnaire may still unintentionally prime respondents or introduce subtle biases, especially when questions involve sensitive or polarizing topics. Although the pollster tries to avoid attrition bias by keeping each interview less than 15 minutes, in seeking to make questions clear and concise, the questionnaire may not capture the full nuance of voter opinions, especially on multifaceted issues like economic policy or social justice. While the poll strives to capture real views without introducing new ideas, respondents are still prone to a social desirability bias. This is especially possible in phone interviews, where respondents may be more inclined to provide answers they perceive as socially acceptable rather than expressing their true opinions. Although the design aims to be fair, the influence of an interviewer’s presence may subtly affect responses (@thepsygologyofsurveyresponse)

\newpage

# Appendix B {-}

# Idealized Methodology

Pursuant to accurately predicting the winner of the presidential election with the electoral college system, there are assumptions that need to be assessed. The United States of America’s electoral college system creates a special environment where votes in some states are disproportionately more influential than that of others and candidates are not elected based on popular vote, but on the majority of electoral votes they get.

## Assumption 1:

Nation-level Popular Vote is irrelevant in our methodology since president electors are chosen based on electoral votes.

##  Assumption 2:

Based on historical and polling data, some states’ electoral votes are predetermined as either Democrat or Republican. Therefore, it is more prudent to shift focus over to swing states where outcomes are uncertain. This assumption is accompanied with the challenge of determining what states qualify as “swing states”. However, on this issue, there is consensus on which states qualify as swing states. Pollsters such as  New York Times have their metric for determining which  states are 'battleground states'. As of 2024, our methodology should direct all our resources to acquiring voting data on the following states (New York Times). This assumption results in each presidential candidate starting off with a set number of electoral votes.

##  Swing States

According to 2024 polling data conducted by New York Times, the swing states are determined to be the following:

Pennsylvania
Georgia
North Carolina
Michigan
Arizona
Wisconsin
Nevada

## Feasibility of Reaching Voters: Justification of Online Panels

According to the New York Times, with a sample of 1000 individual votes, there is a margin of error of around three to four percentages. With $100,000, there are limitations on the labor and resources that is required to poll 1000 voters while balancing out demographics through dialing. Furthermore, there is additional filtering of our sample frame to determine likely voters. Assuming New York Times’ 2% voter response rate, to sample 1000 voters, at least 50,000 voters will need to be called. It is likely that our team will be dialing abundantly more than 50,000 times since it is an unreasonable assumption that every voter we call would pick up or be available to provide a response. There is also the fact that our team does not carry the same credibility and history of acquiring poll data as the New York Times does, resulting in more reluctance for voters to share information with us.
Therefore, it is not feasible for our methodology to employ telephone dialing as our method to reach voters. Instead, we focus on online panels and surveys with incentives for completing the survey correctly. However, online panels and surveys come with their own set of challenges and weaknesses that must be addressed in our methodology. Primarily, the problem of self-selection in online panels disrupts the random selection that allows the external validity of our predictions. Moreover, there is apossibility that we acquire responses from respondents that are not likely to vote, gaining data from outside our target population. Even if we distribute the panels randomly among swing states, responses are still corrupted by selection bias that occurs from responses that are provided from demographics that feel more passionately about the election or are more likely to respond to online panels. Unlike telephone dialing, we have less control over the specific demographic we want to target, resulting in a less stratified data that is representative of the target population overall.

##	Remedies for Online Panels

### Weighting Adjustments

Although online panels offer an imperfect solution to the limited resources our team will work with, there are remedies that can address underrepresentation brought by selection bias. As NYC and Siena College does, post-survey weighting can help us represent our target population more accurately. This is an effective way to adjust our sample based on demographic data and historical turnout patterns within each state. By weighting responses to mirror actual voter demographics from a particular state, the survey results can approximate a more realistic projection of the electorate. These weighting adjustments help to maintain the integrity of our sample, closely resembling the target population of a state.

### Screening for Likely Voters

Although weighting adjustments helpsus represent our target population more closely, there is nothing to stop respondents from outside our target population to participate in our panels since it will be advertised online. To produce accurate predictions of election results, our sample needs to be acquired from a target population that will participate in the election. Therefore, it becomes necessary to include screening questions designed to filter for likely voters, based on voting history or engagement levels. Our survey includes questions that pertain to voting history, interest and engagement, voting intention, and eligibility. Responses to these questions will be used to produce an aggregate “likely voter” score that filters for the target population enabling for more accurate predictions of the presidential election.

##  Elements of Our Survey
To gather accurate responses from a diverse and representative set of likely voters, our survey design carefully adheres to best practices in methodology. Given our goal to avoid attrition bias, we’ve restricted the estimated survey completion time to under ten minutes. This constraint is stricter than the fifteen-minute benchmark typically used by NYC and Siena College for telephone surveys, as online surveys require shorter durations to maintain respondent engagement. Our survey is divided into three main sections: Preamble, Eligibility, and Party Affiliation.

### Preamble
The Preamble serves as an introductory section that aims to establish transparency and motivate respondents to complete the survey accurately. This introduction begins with an explanation of the lottery-style incentive offered for survey completion. Rather than small individual payments, participants have a chance to win a larger prize if they complete the survey, a choice that aligns with budget constraints while enhancing participation rates. To encourage honest responses, respondents are assured of the anonymity of their answers, as well as the fact that their responses will be used solely for research purposes. This transparency fosters trust and can reduce hesitation when sharing political preferences. Information about the survey’s purpose is provided sparingly to minimize any potential response bias, as we communicate that the survey concerns voting and political engagement while omitting specific details that could influence responses, such as particular issues or candidate names. Additionally, respondents have the option to input a referral code if invited by a friend or family member, which can increase their chances of winning the lottery and encourage participation through word-of-mouth. We also request information on the relationship with the referrer to monitor potential demographic clustering in cases where social or familial ties might influence responses.

### Eligibility
The Eligibility section focuses on ensuring that respondents meet the criteria necessary for inclusion in our target population, specifically targeting likely voters within designated swing states to improve the relevance and accuracy of our sample for election predictions. This section begins by asking for respondents' state of residence, age, race, and other preliminary demographic information. To further narrow our sample to those most likely to vote, we employ several screening questions related to voting history, political engagement, and voting intentions. These questions help exclude individuals who are unlikely to participate in the election, thus refining our sample to reflect the likely voting population.

###  Party Affiliation
Party Affiliation addresses respondents’ political leanings and past voting behavior. This section explores their current preference for presidential candidates and parties and asks how they voted in previous elections. By differentiating between party and candidate preference, we can identify any variance in their voting behavior and assess whether respondents’ decisions align more strongly with party or individual candidate considerations. This section also explores whether respondents are informed about their choices, asking questions to determine whether they are aware of both candidates or have researched their policies. By gathering these insights, we can create a profile of the respondent’s political orientation, which allows for a nuanced interpretation of voting intentions that goes beyond simple party preference.

##  Distribution of Online Panels
The online panel method provides valuable flexibility in how surveys are distributed, enabling targeted outreach in regions crucial to our election prediction model. Since the distribution strategy will absorb a significant portion of the budget, this section outlines our plan for cost-effective, targeted, and controlled distribution, aimed at achieving a sample that represents voters in key swing states while addressing biases inherent to online panel surveys.

###   Swing State Representation
To capture a meaningful sample for predicting election results, we target respondents exclusively from key swing states where voting outcomes remain uncertain. By focusing our resources on swing states, we enhance the relevance of our sample to the Electoral College system, which relies heavily on swing state outcomes. Achieving this requires not only careful selection of digital platforms but also strategic advertising in locally relevant spaces.
The lottery-style incentive, with a $5,000 voucher offered to one lucky respondent, serves as an appealing motivator designed to increase participation. We will advertise the survey on websites, forums, and marketplaces that have high traffic within our target states. Examples include social media channels, local news sites, and state-specific online communities. Additionally, we may utilize geo-targeted ads on platforms like Google and Facebook to directly reach residents of these swing states, optimizing our budget by focusing on the digital spaces our target demographic frequents. A significant portion of the budget will thus be allocated to these state-specific online advertisements, which not only promote the survey but also help ensure that respondents fit our target profile.

###   Referral Mechanics
To further broaden our reach without exponentially increasing advertising costs, we incorporate a referral system that rewards respondents for inviting others to complete the survey. Each referral allows the original respondent to gain an additional entry in the lottery, increasing their chances of winning. This referral incentive taps into snowball sampling, where respondents refer people within their network to participate, expanding our pool of respondents organically.
However, using snowball sampling in an election survey introduces potential biases, particularly the risk of overrepresentation of individuals with similar political views, as respondents are likely to refer to close contacts who may share similar ideologies. To mitigate this, we will place specific restrictions on the referral program. Each respondent will be limited to referring only one other individual, minimizing the potential for overly clustered referrals. Additionally, we will require respondents to identify their relationship with each referral, helping us monitor homogeneity in our responses. By balancing the benefits of snowball sampling with these restrictions, we aim to expand the reach of our survey into difficult to reach demographics without compromising the representativeness of the sample.

# Budget Allocation
  Advertisements for sampling recruitment: $70,000
  Prize Pool for Incentive: $5,000
  Data Processing and Cleaning: $15,000
  Miscellaneous Expenses: $5,000
  Petty Cash: $5,000

#  Idealized Questionnaire and Survey

https://docs.google.com/forms/d/117wC8f1rBW_bd5glEdiR-7823-02A3HiD5EzQuT1GVw/edit

# Structure

## Preamble

Title: 2024 Presidential Election Poll Data

Introduction: Thank you for your participation in our $5,000 giveaway program.

For entry, please complete the survey in its entirety below. Please be informed that your answers are strictly confidential and will, under no circumstances, be shared with any party, group, or persons of affiliation.

$5,000 Cash Voucher Program Details: Completing the survey below qualifies as one entry into our giveaway. Each participant receives a unique entry identifier, and every entry will have an equal chance in our final random selection of the prize winner. Please note that each individual may complete only one survey.

Referral Program: Our referral program offers you the chance to increase your odds of winning. After completing the survey, you’ll receive a unique referral code. If a friend or associate completes the survey using your code, your total number of entries doubles, effectively doubling your chances of winning the $5,000 prize.

Finally, we wish to clarify that this survey is conducted purely for research purposes and is entirely unaffiliated with any political party, organization, or individual in a position of power. Our goal is to gain insight into the evolving political landscape and levels of public engagement.

Referral Code (please ignore if inapplicable)

If you have entered a referral code above, please indicate the relationship with the referee from the best choice below.
Friend
Associate
Relative
Other…

## Eligibility
Do you meet your state's residency requirements?
Yes
No (End Survey)
Please select "N/A" from the options below if you reside in North Dakota. 
Which of the following best describes you?
I am registered to vote
I am not registered to vote
I plan to register
N/A
Full Name (For identification of $5,000 cash winner)
From which state will you be vote for the next president of the United States of America?
Where can we reach you when we announce the winner?
E-mail
Telephone
Other…
Contact information
**We will not use this for any other purpose than to announce winners. It will only be used to reach our lucky winner for instructions on receiving the cash prize.
Which of the following age groups best describes you?
18-25
26-31
32-40
41-50
51-60
61-80
81-125
Under 18
Which of the following best describes you?
American Indian or Alaskan Native
Asian
African American
Hispanic or Latino
Native Hawaiian or Pacific Islander
White
Middle Eastern or North African
Other…
What is your gender?
Male
Female
Prefer not to say
Other…
Which of the following situations best describe you?
I did not vote in the 2020 presidential election
I planned to vote but wasn’t able to
I voted in the 2020 presidential election
How likely are you to vote in the upcoming election?
Very unlikely
1
2
3
Very likely
Have you attended a presidential rally or their campaigns this year?
Yes
No
Only in the past
On a scale of 1 to 3, how closely do you follow the presidential race?
Not at all
1
2
3
4
5
Very closely

## Party Affiliation
Which party do you prefer a candidate to be chosen from?
Democratic Party's candidate
Republican Party's candidate
Another Party's candidate
In the 2020 presidential election, who did you vote for?  If you did not vote, who would you have voted for?
Donald Trump
Joe Biden
Neither
Other…
If planning to vote, which presidential candidate are you planning to vote for?
If you are abstaining, please select which candidate you feel more strongly about.
Donald Trump
Kamala Harris
Other...
\newpage

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

```{r}
#| echo: false
#| warning: false
#| message: false
posterior_predictions <- posterior_predict(model, draws = 100)
ppc_dens_overlay(y = data_filtered$average_pct/100, yrep = posterior_predictions) +
  labs(title = "Posterior Predictive Check",
       x = "Rate of support",
       y = "Density")
```

## Diagnostics

```{r}

```

\newpage

# References