### 2024-11-04 paper/paper.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Forecasting the 2024 U.S. Presidential Election},
  pdfauthor={Andy Jiang; Onon Burentuvshin; Rui Hu},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Forecasting the 2024 U.S. Presidential Election\thanks{Code and
data are available at:
{[}https://github.com/AndyYanxunJiang/president-polls).}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Using Aggregated Polling Data to Predict the Outcome of the
2024 Presidential Race}
\author{Andy Jiang \and Onon Burentuvshin \and Rui Hu}
\date{November 4, 2024}

\begin{document}
\maketitle
\begin{abstract}
The 2024 U.S. presidential election is shaping up to be a highly swing
race, with polling data playing a crucial role in gauging public
sentiment and forecasting the potential outcome. This paper aims to
build a predictive model using a poll-of-polls approach, aggregating
data from multiple pollsters to forecast the winner of the upcoming
election. The analysis incorporates polling data from various sources,
filtered to include only high-quality pollsters with a numeric grade
above 3, to ensure the reliability of the predictions. Additionally, a
detailed examination of one selected pollster's methodology is
conducted, highlighting strengths, weaknesses, and biases in the survey
process. The paper concludes with an idealized polling methodology,
designed within a \$100K budget, to enhance the accuracy of election
forecasts. Results are presented in the context of understanding polling
trends, potential biases, and the evolving landscape of voter
preferences as the election approaches.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden, boxrule=0pt]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The 2024 U.S. presidential election is shaping up to be one of the most
closely watched and potentially pivotal elections in recent history. As
public opinion shifts and political dynamics evolve, polling data has
emerged as a critical tool for forecasting the outcome and understanding
voter preferences. Polls provide a snapshot of where candidates stand at
various points in time, reflecting the impact of campaigns, debates, and
major events. However, the reliability and accuracy of individual polls
can vary significantly based on methodology, sample size, and other
factors, making it necessary to aggregate multiple sources of data for a
more robust prediction.

This paper aims to forecast the outcome of the 2024 U.S. presidential
election by leveraging a ``poll-of-polls'' approach, which aggregates
data from various pollsters to create a more comprehensive picture of
the race. By focusing on high-quality polls and conducting a detailed
analysis of the polling trends for the two main candidates, Donald Trump
and Kamala Harris, we seek to uncover patterns in voter support and
potential factors influencing shifts in public opinion. Additionally,
the paper examines the methodologies of different pollsters to assess
their strengths and weaknesses, which can help identify possible biases
in the aggregated data.

The analysis will further explore the geographic distribution of support
across key swing states, providing insights into the electoral
landscape. Finally, the paper presents an idealized polling methodology
with a \$100,000 budget, designed to improve the accuracy and
representativeness of election forecasts. By integrating multiple data
sources and critically evaluating polling practices, this study aims to
contribute to the ongoing discussion on the role of polling in
democratic processes and election forecasting.

\hypertarget{sec-data}{%
\section{Data}\label{sec-data}}

The data used in this paper was obtained from the FiveThirtyEight
website (Best, Bycoffe, et al. 2024), which aggregates polling data and
other relevant indicators to track the U.S. election landscape. We
accessed this data using custom web scraping tools and APIs provided by
FiveThirtyEight. The dataset includes information on pollster, state,
start and end dates of polls, sample size, population type (e.g.,
registered voters or likely voters), and party affiliation.

\hypertarget{overview}{%
\subsection{Overview}\label{overview}}

Our analysis was conducted using R (R Core Team 2022) along with several
supporting packages. We used the \textbf{tidyverse} for data
manipulation (Wickham et al. 2019), \textbf{readr} for data import
(Wickham, Hester, and Bryan 2023), and \textbf{here} for file management
(Müller 2020).

Dynamic reporting was handled with \textbf{knitr} (Xie 2024), while
\textbf{kableExtra} allowed for enhanced table formatting (Zhu 2024). We
used \textbf{scales} to adjust visualization scales (Wickham, Pedersen,
and Seidel 2023), and \textbf{arrow} for efficient data processing
(Richardson et al. 2024).

For Bayesian modeling and visualization, we used \textbf{bayesplot}
(Gabry and Mahr 2024) and \textbf{rstanarm} (Goodrich et al. 2024).

\hypertarget{measurement}{%
\subsection{Measurement}\label{measurement}}

The FiveThirtyEight dataset is updated in real-time, providing the
latest polling data for the 2024 U.S. presidential election. This
dataset aggregates information from multiple reputable polling sources,
applying adjustments to enhance reliability, such as weighting based on
pollster quality and recency of data. These measures ensure that the
dataset remains a dependable reflection of current voter sentiment and
trends as Election Day approaches.

\hypertarget{analysis-data}{%
\subsection{Analysis Data}\label{analysis-data}}

The analysis dataset includes several key variables: State, Sample Size,
Population Type (e.g., Registered Voters), Party Affiliation, Candidate
Preference (Trump or Harris), and Pollster. These variables allow for an
analysis of voter sentiment in specific swing states, with a focus on
support levels between Donald Trump and Kamala Harris.

Certain variables, such as Question ID, Start Date, End Date, and
Hypothetical Questions, were removed as they were not necessary for this
analysis. Additionally, data on other candidates was excluded to
concentrate solely on the Trump vs.~Harris matchup, aligning with the
focus on the most relevant contenders in the seven swing states:
Pennsylvania, Georgia, North Carolina, Michigan, Arizona, Wisconsin,
Nevada.

Table~\ref{tbl-sample} provides a random sample of the analysis dataset,
displaying all relevant variables used for analysis.

\hypertarget{tbl-sample}{}
\begin{longtable}[t]{>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{2.5cm}>{}r>{\raggedright\arraybackslash}p{1.5cm}>{\raggedright\arraybackslash}p{1.5cm}ll}
\caption{\label{tbl-sample}Random sample of 10 rows from the analysis data. }Sample of 10 Random Observations from the Analysis Data}\\
\toprule
pollster & state & sample\_size & population & party & answer & candidate\_name\\
\midrule
PPP & Arizona & 596 & rv & REP & Trump & Donald Trump\\
Morning Consult & Wisconsin & 700 & rv & REP & Trump & Donald Trump\\
Morning Consult & Georgia & 800 & rv & REP & Trump & Donald Trump\\
Morning Consult & Nevada & 574 & rv & REP & Trump & Donald Trump\\
Bullfinch & Pennsylvania & 800 & rv & DEM & Harris & Kamala Harris\\
\addlinespace
Public Opinion Strategies & Pennsylvania & 500 & rv & REP & Trump & Donald Trump\\
High Point University & North Carolina & 804 & rv & REP & Trump & Donald Trump\\
Emerson & Michigan & 1000 & rv & REP & Trump & Donald Trump\\
Quinnipiac & Georgia & 1203 & rv & REP & Trump & Donald Trump\\
Emerson & Pennsylvania & 850 & rv & REP & Trump & Donald Trump\\
\bottomrule
\end{longtable}

\hypertarget{predictor-variables}{%
\subsection{Predictor variables}\label{predictor-variables}}

\hypertarget{model}{%
\section{Model}\label{model}}

In this analysis, we utilize a logistic regression model to predict
polling percentages for the 2024 U.S. presidential candidates over time.
In order to focus on states with historical significance in elections.
The model is specified as a Generalized Linear Model (GLM) with a
binomial family and a logit link function, appropriate for modeling
binary or proportion-based outcomes.

\hypertarget{model-set-up}{%
\subsection{Model set-up}\label{model-set-up}}

The dataset, model\_data.csv, provides polling data with average\_pct as
the observed average polling percentage among 7 swing states. The
end\_date indicates the polling period's endpoint. To predict polling
percentages beyond the available data, we created an additional sequence
of 20 days beyond the dataset's last recorded date. Because we want to
predict the result in November 5. This expansion allows for forecasting
near-term polling trends. Also, capturing potential shifts in candidate
support leading up to election day. The model was trained on existing
data and applied to this extended date range using predict(model,
newdata = ., type = ``response'').

Our model setup and variable selection are as follows: \[
\text{logit}(p_i) = \beta_0 + \beta_1 (\text{Date}) + \beta_2 (\text{Candidate})
\]

where: -
\[p_i \ \text{represents the probability of support for candidate i at a specific date.}\]
- \[\beta_0 \ \text{is the intercept, representing baseline support.}\]
-
\[\beta_1 \ \text{captures changes in support over time (date effect).}\]
-
\[\beta_2 \ \text{is a categorical variable accounting for each candidate.}\]

\hypertarget{model-justification}{%
\subsubsection{Model justification}\label{model-justification}}

The choice of a logistic regression model was guided by the nature of
polling data. Nature of polling data represents proportions that
naturally bound between 0 and 1. The logistic function's S-shaped curve
effectively models the relationship between time and polling percentage,
accommodating non-linear trends that may arise as election day nears.
Firstly, candidate names should be included as a categorical predictor
in the model. It allows us to capture individual impact in support rates
by candidate. In addition, a transformation from date to numeric is made
on end\_date in order to get a numerical format. It is called the
end\_date\_numeric. It helps put date as a predictor in the model. By
structuring the model this way, it can predict the percentage of the
poll by both the impact of candidates and trends over time.

\hypertarget{results}{%
\section{Results}\label{results}}

The analysis of polling data across key swing states for the 2024 U.S.
Presidential Election reveals insights into the levels of support for
candidates Donald Trump and Kamala Harris. The findings are presented
through three primary visualizations: the average polling percentage by
state, the distribution of polling percentages, and the difference in
average polling percentage between the two candidates.

\hypertarget{average-polling-percentage-by-state}{%
\subsection{Average Polling Percentage by
State}\label{average-polling-percentage-by-state}}

From Figure~\ref{fig-1}, we can see that in the swing states of Arizona,
Georgia, and North Carolina, Trump holds a narrow lead. In contrast,
Harris leads in Pennsylvania, Wisconsin, Nevada, and Michigan. When
looking at the overall average across these states, Harris slightly
edges out Trump, indicating her overall advantage in these swing areas.

\hypertarget{polling-percentage-distribution}{%
\subsection{Polling Percentage
Distribution}\label{polling-percentage-distribution}}

The box plot in Figure~\ref{fig-2} demonstrates the range and
consistency of support for both candidates across each state. Trump
shows a wider range of support, reflecting variability, while Harris has
relatively steady polling percentages across states, maintaining higher
consistency. Harris's support is less variable and generally centers
around higher polling percentages compared to Trump's.

\hypertarget{difference-in-average-polling-percentage}{%
\subsection{Difference in Average Polling
Percentage}\label{difference-in-average-polling-percentage}}

Figure~\ref{fig-3} highlights the difference in average polling
percentages by state, with positive values indicating Trump's lead and
negative values indicating Harris's. Trump's lead in the three states
(Arizona, Georgia, and North Carolina) is relatively narrow, with
margins of less than 1 percent. In contrast, Harris's lead in the states
favoring her (Pennsylvania, Wisconsin, Nevada, and Michigan) ranges from
1 to 4 percent, indicating her stronger support in the states where she
leads.

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-1-1.pdf}

}

\caption{\label{fig-1}Average Polling Percentage by State}

\end{figure}

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-2-1.pdf}

}

\caption{\label{fig-2}Polling Percentage Distribution}

\end{figure}

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-3-1.pdf}

}

\caption{\label{fig-3}Difference in Average Polling Percentage}

\end{figure}

\hypertarget{swing-state-polling-prediction-model}{%
\subsection{Swing State Polling Prediction
Model}\label{swing-state-polling-prediction-model}}

Figure~\ref{fig-4} shows the trend of support rate changes for each
candidate before the upcoming election, as well as the model's predicted
values for the next 20 days.We predicts that Donald Trump may win based
on our model.

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-4-1.pdf}

}

\caption{\label{fig-4}Swing State Polling Prediction Model}

\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{impact-of-key-demographics-and-policy-appeals}{%
\subsection{Impact of Key Demographics and Policy
Appeals}\label{impact-of-key-demographics-and-policy-appeals}}

Harris's lead in several swing states may be influenced by her
campaign's appeal to key demographic groups, particularly immigrants.
The Democratic platform has emphasized social and economic benefits for
immigrant communities, which may have strengthened her base in states
with significant immigrant populations. Policies that promise support
for healthcare, education, and other public services are likely
contributing to Harris's appeal among these groups, solidifying her lead
in states like Nevada and Michigan. This demographic support provides
Harris with a strategic advantage, as these communities may play a
crucial role in voter turnout and loyalty.

\hypertarget{influence-of-endorsements-and-high-profile-backing}{%
\subsection{Influence of Endorsements and High-Profile
Backing}\label{influence-of-endorsements-and-high-profile-backing}}

Recently, high-profile figures like Elon Musk have publicly thrown their
support behind Trump, leveraging their influence to rally support for
his campaign (Zahn 2024). Musk's endorsement could invigorate Trump's
base, especially among Musk's followers and those engaged in the fields
of business and technology. This visible backing from such influential
figures has the potential to sway undecided voters and influence
last-minute decisions, particularly in swing states where Trump's lead
remains narrow.

Musk's endorsement might not only energize Trump's existing supporters
but could also lead to a boost in campaign funds and social media
engagement, further strengthening Trump's position. In closely contested
states, where a few thousand votes can make a significant difference,
this level of backing from a high-profile figure could be crucial. If
Musk's involvement draws in additional resources and amplifies Trump's
visibility, it may ultimately impact the final outcome of the election
on November 5th.

\hypertarget{limitations-of-the-analysis-focus-on-swing-states}{%
\subsection{Limitations of the Analysis: Focus on Swing
States}\label{limitations-of-the-analysis-focus-on-swing-states}}

While the analysis focuses on key swing states, it does not consider the
broader national picture, which includes states with clear, one-sided
advantages for either candidate. This narrowed focus on swing states
provides insights into swing areas but does not give a complete picture
of the national landscape. Therefore, while Harris appears to have an
edge in these swing states, it is unclear whether she is leading
overall. The situation in solidly red or blue states could still shift
the electoral balance, meaning Trump could potentially gain ground
through strong performances in traditionally Republican states.

\hypertarget{potential-influence-of-late-campaign-efforts}{%
\subsection{Potential Influence of Late Campaign
Efforts}\label{potential-influence-of-late-campaign-efforts}}

With a few days remaining before the election, both campaigns are likely
to intensify their outreach efforts, particularly in swing states.
Trump's recent increase in targeted advertisements and campaign rallies
in swing states could energize his supporters, narrowing Harris's lead.
On the other hand, Harris's campaign may focus on mobilizing key
demographics, such as young and minority voters, to secure her
advantage. Late-breaking events, public debates, or controversies could
sway undecided voters and ultimately impact the election results in
these states.

\hypertarget{limitations-and-next-steps}{%
\subsection{Limitations and Next
Steps}\label{limitations-and-next-steps}}

This analysis is limited to examining current polling data in swing
states and does not account for unexpected shifts in voter sentiment or
turnout on election day. Additionally, focusing exclusively on swing
states omits the influence of non-swing states that may still play a
role in the electoral vote tally. Future analyses could incorporate
national polling data to provide a more comprehensive perspective and
examine demographic breakdowns within each state to understand which
groups are driving support for each candidate. Integrating data on
recent campaign activities, endorsements, and voter sentiment on
pressing issues could also improve the predictive accuracy of election
outcomes.

\newpage

\appendix

\hypertarget{appendix-a}{%
\section*{Appendix A}\label{appendix-a}}
\addcontentsline{toc}{section}{Appendix A}

\hypertarget{sampling-methodology-employed-by-new-york-times-and-siena-college}{%
\section{Sampling Methodology Employed by New York Times and Siena
College:}\label{sampling-methodology-employed-by-new-york-times-and-siena-college}}

Our analysis and prediction utilizes data collected by a single pollster
and it is important to scrutinize the sampling methodology behind the
sampling methods used to determine the extent of the external validity
of our predictions. Primarily, the data gathered by New York Times was
done so with the collaboration with Sienna College through
live-interviewers reaching voters by telephone.

\hypertarget{target-population-and-sampling-frame}{%
\subsection{Target Population and Sampling
Frame}\label{target-population-and-sampling-frame}}

The poll targets registered voters in the United States, with a
particular focus on voters in swing states for presidential races.
Consequently, the target population is registered voters of the United
States of America and the sampling frame is the voter file, a national
list of registered voters from which NYC employs random sampling to
create their pollsters The New York Times (2024). This voter file
contains information on more than 200 million Americans that includes
information about their party affiliation, voting history, and basic
demographic information. From this sampling frame, respondents are
randomly selected while balancing key demographic and political
characteristics through both front-end sampling and post-survey
weighting pursuant to appropriate representation of the entire target
population The New York Times (2024).

\hypertarget{recruitment-methods}{%
\subsection{Recruitment Methods}\label{recruitment-methods}}

The poll relies on random digit dialing to reach voters, predominantly
via cell phones. While landlines are also called, the dominance of
cellphone usage aligns with contemporary communication trends. This
results in their rendition of random sampling, which contributes to the
external validity of data. However, reaching voters by phone has become
more difficult due to low response rates requiring many attempts to
secure responses from certain individuals (response rates are as low as
2\% at NYC). In fact, according to an article by Emerson College,
telephone polling has become entirely obsolete Kimball, Mumford, and
Taglia (2024). Despite this, the phone methodology is still seen as
effective for its ability to quickly reach a diverse voter pool across
various states as well as seen as the golden standard for polling.
Although this can be seen as random sampling, the concern of low
response rates and reliance on dialing to reach voters raises concerns
of selection bias. There are questions as to whether a certain
demographic will be more responsive to being reached through a cell
phone. Moreover, voters that are reached over the phone may feel the
urge to provide more `socially desirable' responses that are not
reflective of their true opinions.

\hypertarget{non-response-remedies}{%
\subsection{Non-Response Remedies}\label{non-response-remedies}}

With a response rate of 2\%, there is valid concern over non-response
bias underrepresenting certain demographics. The poll addresses
non-response through a combination of front-end sampling adjustments
(calling harder-to-reach voters) and post-survey weighting. Weighting is
applied to ensure that the final sample aligns with the demographic and
political composition of the electorate, based on voter file data.
Non-response bias is a persistent challenge, and even though weighting
helps, it's impossible to be certain that respondents are fully
representative of those who do not respond. Experiments with alternative
methods, such as mail surveys with incentives of cash vouchers have
shown to increase response rate up to 30\% The New York Times (2024).

\hypertarget{sampling-approach-stratified-sampling}{%
\section{Sampling Approach: Stratified
Sampling}\label{sampling-approach-stratified-sampling}}

\hypertarget{methodology}{%
\subsection{Methodology}\label{methodology}}

The Times/Siena Poll employs a stratified random sampling approach to
ensure that the sample accurately represents the diversity within the
target population of registered voters in the United States. This method
involves dividing the population into distinct subgroups, or ``strata,''
based on key characteristics such as state, age, gender, race, education
level, and political affiliation. For example, in presidential
elections, particular emphasis is placed on voters from swing states
like Arizona, North Carolina, Georgia, and Pennsylvania. By focusing on
these battleground states, the poll aims to capture the opinions of
voters who are most likely to influence the election outcome. Once the
strata are established, respondents are randomly selected within each
group, ensuring that all demographic groups are appropriately
represented in proportion to their actual distribution within the
electorate. Furthermore, post-survey weighting is employed to more
closely resemble this distribution.

\hypertarget{advantages}{%
\subsection{Advantages}\label{advantages}}

Stratified sampling offers the advantage of higher precision in
representing the population's diversity, as each subgroup is
proportionally represented. By breaking down the sample into distinct
groups, the Times/Siena Poll can adjust for demographic variations
between states or voter groups, which is particularly useful in a
diverse electorate. This approach helps to minimize sampling error by
ensuring that all significant voter subgroups are included, thus
increasing the reliability of the poll's predictions. It is standard
practice for them to sample 1,000 respondents for their polls. Using
stratified sampling to have this sample size closely resemble that of
the entire target population cuts costs and provides efficient real-time
data.

\hypertarget{stratified-sampling-over-random-sampling}{%
\subsection{Stratified Sampling over Random
Sampling}\label{stratified-sampling-over-random-sampling}}

If the methodology randomly sampled from the target population without
regard for the distribution of demographics in the state, significant
increases to the sample sizes have to be made. Random sampling will have
largely less precision in representing the distribution of different
demographics within a given state unless a large enough sample size were
to be drawn. Although it is ideal to have probability based sampling
increasingly resemble the demographic distribution quota with growing
sample sizes, it is unfeasible to conduct such a large scale poll as
this approach would encounter a limitation of resources. With declining
response rates and the difficulty of reaching voters, New York
Times/Siena's nonprobability sampling is a more realistic and efficient
method of polling than probability sampling is
(\textbf{tellingstories?}). As such, the stratified sampling approach
allows New York Times/Siena to represent the entire target population
with a sample size of 1,000.

\hypertarget{challenges-and-trade-offs}{%
\subsection{Challenges and Trade-offs}\label{challenges-and-trade-offs}}

Although stratified sampling improves representativeness, it presents
certain challenges in execution. First, identifying and reaching a
sufficient number of respondents from harder-to-reach groups (such as
young voters, minority populations, and those without reliable phone
access) can be difficult, especially given the 2\% response rate. To
counterbalance the low response rates, the poll requires multiple call
attempts and careful post-survey weighting to adjust for any potential
underrepresentation. Additionally, while phone polling offers broad
geographic reach and a standardized mode of data collection, it limits
the sample to voters with phone access who are willing to participate in
phone interviews, potentially introducing selection bias if certain
demographics are less reachable by phone. Although New York Times/Sienna
College employs random digit dialing from their sampling frame, their
rendition of sample recruitment requires them to sample individuals
according to quotas, resulting in nonprobability sampling.

\hypertarget{evaluation-of-questionnaire-design-in-the-timessiena-poll}{%
\subsection{Evaluation of Questionnaire Design in the Times/Siena
Poll}\label{evaluation-of-questionnaire-design-in-the-timessiena-poll}}

\hypertarget{strengths}{%
\subsubsection{Strengths}\label{strengths}}

The Times/Siena Poll's approach to questionnaire design reflects a
careful effort to craft unbiased, relevant, and inclusive questions. By
tailoring questions to current events and topics of widespread interest,
the poll aims to capture public sentiment on issues that directly affect
voters' attitudes, such as economic conditions or recent political
debates. This responsiveness to current events strengthens the poll's
relevance and allows it to track shifts in opinion over time. The poll's
commitment to neutrality is a significant strength. Each question is
designed to ensure that respondents, regardless of their political
views, feel that their perspectives are fairly represented. The team
behind the poll spends substantial time crafting response options to
reflect a balanced spectrum of views, minimizing potential response bias
and enhancing the questionnaire's reliability across different
demographic and ideological groups The New York Times (2024). Clarity is
also a central focus in their questionnaire development. The team
emphasizes ensuring that each question is clear and unambiguous,
minimizing the risk of misinterpretation. This dedication to
straightforward, widely understandable questions improves the validity
of responses, as it reduces the chance that respondents interpret
questions differently based on background or knowledge level.

\hypertarget{limitations}{%
\subsubsection{Limitations}\label{limitations}}

Despite the strengths of the Times/Siena Poll's questionnaire design,
there are some limitations. One notable concern is that, although the
poll tries to avoid leading questions, any standardized questionnaire
may still unintentionally prime respondents or introduce subtle biases,
especially when questions involve sensitive or polarizing topics.
Although the pollster tries to avoid attrition bias by keeping each
interview less than 15 minutes, in seeking to make questions clear and
concise, the questionnaire may not capture the full nuance of voter
opinions, especially on multifaceted issues like economic policy or
social justice. While the poll strives to capture real views without
introducing new ideas, respondents are still prone to a social
desirability bias. This is especially possible in phone interviews,
where respondents may be more inclined to provide answers they perceive
as socially acceptable rather than expressing their true opinions.
Although the design aims to be fair, the influence of an interviewer's
presence may subtly affect responses Roger Tourangeau and Rasinski
(2000).

\newpage

\hypertarget{appendix-b}{%
\section*{Appendix B}\label{appendix-b}}
\addcontentsline{toc}{section}{Appendix B}

Pursuant to accurately predicting the winner of the presidential
election with the electoral college system, there are assumptions that
need to be assessed. The United States of America's electoral college
system creates a special environment where votes in some states are
disproportionately more influential than that of others and candidates
are not elected based on popular vote, but on the majority of electoral
votes they get.

\hypertarget{assumption-1}{%
\subsection{Assumption 1:}\label{assumption-1}}

Nation-level Popular Vote is irrelevant in our methodology since
president electors are chosen based on electoral votes.

\hypertarget{assumption-2}{%
\subsection{Assumption 2:}\label{assumption-2}}

Based on historical and polling data, some states' electoral votes are
predetermined as either Democrat or Republican. Therefore, it is more
prudent to shift focus over to swing states where outcomes are
uncertain. This assumption is accompanied with the challenge of
determining what states qualify as ``swing states''. However, on this
issue, there is consensus on which states qualify as swing states.
Pollsters such as New York Times have their metric for determining which
states are `battleground states'. As of 2024, our methodology should
direct all our resources to acquiring voting data on the following
states (New York Times). This assumption results in each presidential
candidate starting off with a set number of electoral votes.

\hypertarget{swing-states}{%
\subsection{Swing States}\label{swing-states}}

According to 2024 polling data conducted by New York Times, the swing
states are determined to be the following Boyce et al. (2024):

Pennsylvania Georgia North Carolina Michigan Arizona Wisconsin Nevada

\hypertarget{feasibility-of-reaching-voters-justification-of-online-panels}{%
\subsection{Feasibility of Reaching Voters: Justification of Online
Panels}\label{feasibility-of-reaching-voters-justification-of-online-panels}}

According to the New York Times, with a sample of 1000 individual votes,
there is a margin of error of around three to four percentages. With
\$100,000, there are limitations on the labor and resources that is
required to poll 1000 voters while balancing out demographics through
dialing. Furthermore, there is additional filtering of our sample frame
to determine likely voters. Assuming New York Times' 2\% voter response
rate, to sample 1000 voters, at least 50,000 voters will need to be
called. It is likely that our team will be dialing abundantly more than
50,000 times since it is an unreasonable assumption that every voter we
call would pick up or be available to provide a response. There is also
the fact that our team does not carry the same credibility and history
of acquiring poll data as the New York Times does, resulting in more
reluctance for voters to share information with us. Therefore, it is not
feasible for our methodology to employ telephone dialing as our method
to reach voters. Instead, we focus on online panels and surveys with
incentives for completing the survey correctly. However, online panels
and surveys come with their own set of challenges and weaknesses that
must be addressed in our methodology. Primarily, the problem of
self-selection in online panels disrupts the random selection that
allows the external validity of our predictions. Moreover, there is
apossibility that we acquire responses from respondents that are not
likely to vote, gaining data from outside our target population. Even if
we distribute the panels randomly among swing states, responses are
still corrupted by selection bias that occurs from responses that are
provided from demographics that feel more passionately about the
election or are more likely to respond to online panels. Unlike
telephone dialing, we have less control over the specific demographic we
want to target, resulting in a less stratified data that is
representative of the target population overall.

\hypertarget{remedies-for-online-panels}{%
\subsection{Remedies for Online
Panels}\label{remedies-for-online-panels}}

\hypertarget{weighting-adjustments}{%
\subsubsection{Weighting Adjustments}\label{weighting-adjustments}}

Although online panels offer an imperfect solution to the limited
resources our team will work with, there are remedies that can address
underrepresentation brought by selection bias. As NYC and Siena College
does, post-survey weighting can help us represent our target population
more accurately. This is an effective way to adjust our sample based on
demographic data and historical turnout patterns within each state. By
weighting responses to mirror actual voter demographics from a
particular state, the survey results can approximate a more realistic
projection of the electorate. These weighting adjustments help to
maintain the integrity of our sample, closely resembling the target
population of a state.

\hypertarget{screening-for-likely-voters}{%
\subsubsection{Screening for Likely
Voters}\label{screening-for-likely-voters}}

Although weighting adjustments helpsus represent our target population
more closely, there is nothing to stop respondents from outside our
target population to participate in our panels since it will be
advertised online. To produce accurate predictions of election results,
our sample needs to be acquired from a target population that will
participate in the election. Therefore, it becomes necessary to include
screening questions designed to filter for likely voters, based on
voting history or engagement levels. Our survey includes questions that
pertain to voting history, interest and engagement, voting intention,
and eligibility. Responses to these questions will be used to produce an
aggregate ``likely voter'' score that filters for the target population
enabling for more accurate predictions of the presidential election.

\hypertarget{elements-of-our-survey}{%
\subsection{Elements of Our Survey}\label{elements-of-our-survey}}

To gather accurate responses from a diverse and representative set of
likely voters, our survey design carefully adheres to best practices in
methodology. Given our goal to avoid attrition bias, we've restricted
the estimated survey completion time to under ten minutes. This
constraint is stricter than the fifteen-minute benchmark typically used
by NYC and Siena College for telephone surveys, as online surveys
require shorter durations to maintain respondent engagement. Our survey
is divided into three main sections: Preamble, Eligibility, and Party
Affiliation.

\hypertarget{preamble}{%
\subsubsection{Preamble}\label{preamble}}

The Preamble serves as an introductory section that aims to establish
transparency and motivate respondents to complete the survey accurately.
This introduction begins with an explanation of the lottery-style
incentive offered for survey completion. Rather than small individual
payments, participants have a chance to win a larger prize if they
complete the survey, a choice that aligns with budget constraints while
enhancing participation rates. To encourage honest responses,
respondents are assured of the anonymity of their answers, as well as
the fact that their responses will be used solely for research purposes.
This transparency fosters trust and can reduce hesitation when sharing
political preferences. Information about the survey's purpose is
provided sparingly to minimize any potential response bias, as we
communicate that the survey concerns voting and political engagement
while omitting specific details that could influence responses, such as
particular issues or candidate names. Additionally, respondents have the
option to input a referral code if invited by a friend or family member,
which can increase their chances of winning the lottery and encourage
participation through word-of-mouth. We also request information on the
relationship with the referrer to monitor potential demographic
clustering in cases where social or familial ties might influence
responses.

\hypertarget{eligibility}{%
\subsubsection{Eligibility}\label{eligibility}}

The Eligibility section focuses on ensuring that respondents meet the
criteria necessary for inclusion in our target population, specifically
targeting likely voters within designated swing states to improve the
relevance and accuracy of our sample for election predictions. This
section begins by asking for respondents' state of residence, age, race,
and other preliminary demographic information. To further narrow our
sample to those most likely to vote, we employ several screening
questions related to voting history, political engagement, and voting
intentions. These questions help exclude individuals who are unlikely to
participate in the election, thus refining our sample to reflect the
likely voting population.

\hypertarget{party-affiliation}{%
\subsubsection{Party Affiliation}\label{party-affiliation}}

Party Affiliation addresses respondents' political leanings and past
voting behavior. This section explores their current preference for
presidential candidates and parties and asks how they voted in previous
elections. By differentiating between party and candidate preference, we
can identify any variance in their voting behavior and assess whether
respondents' decisions align more strongly with party or individual
candidate considerations. This section also explores whether respondents
are informed about their choices, asking questions to determine whether
they are aware of both candidates or have researched their policies. By
gathering these insights, we can create a profile of the respondent's
political orientation, which allows for a nuanced interpretation of
voting intentions that goes beyond simple party preference.

\hypertarget{distribution-of-online-panels}{%
\subsection{Distribution of Online
Panels}\label{distribution-of-online-panels}}

The online panel method provides valuable flexibility in how surveys are
distributed, enabling targeted outreach in regions crucial to our
election prediction model. Since the distribution strategy will absorb a
significant portion of the budget, this section outlines our plan for
cost-effective, targeted, and controlled distribution, aimed at
achieving a sample that represents voters in key swing states while
addressing biases inherent to online panel surveys.

\hypertarget{swing-state-representation}{%
\subsubsection{Swing State
Representation}\label{swing-state-representation}}

To capture a meaningful sample for predicting election results, we
target respondents exclusively from key swing states where voting
outcomes remain uncertain. By focusing our resources on swing states, we
enhance the relevance of our sample to the Electoral College system,
which relies heavily on swing state outcomes. Achieving this requires
not only careful selection of digital platforms but also strategic
advertising in locally relevant spaces. The lottery-style incentive,
with a \$5,000 voucher offered to one lucky respondent, serves as an
appealing motivator designed to increase participation. We will
advertise the survey on websites, forums, and marketplaces that have
high traffic within our target states. Examples include social media
channels, local news sites, and state-specific online communities.
Additionally, we may utilize geo-targeted ads on platforms like Google
and Facebook to directly reach residents of these swing states,
optimizing our budget by focusing on the digital spaces our target
demographic frequents. A significant portion of the budget will thus be
allocated to these state-specific online advertisements, which not only
promote the survey but also help ensure that respondents fit our target
profile.

\hypertarget{referral-mechanics}{%
\subsubsection{Referral Mechanics}\label{referral-mechanics}}

To further broaden our reach without exponentially increasing
advertising costs, we incorporate a referral system that rewards
respondents for inviting others to complete the survey. Each referral
allows the original respondent to gain an additional entry in the
lottery, increasing their chances of winning. This referral incentive
taps into snowball sampling, where respondents refer people within their
network to participate, expanding our pool of respondents organically.
However, using snowball sampling in an election survey introduces
potential biases, particularly the risk of overrepresentation of
individuals with similar political views, as respondents are likely to
refer to close contacts who may share similar ideologies. To mitigate
this, we will place specific restrictions on the referral program. Each
respondent will be limited to referring only one other individual,
minimizing the potential for overly clustered referrals. Additionally,
we will require respondents to identify their relationship with each
referral, helping us monitor homogeneity in our responses. By balancing
the benefits of snowball sampling with these restrictions, we aim to
expand the reach of our survey into difficult to reach demographics
without compromising the representativeness of the sample.

\hypertarget{budget-allocation}{%
\section{Budget Allocation}\label{budget-allocation}}

Advertisements for sampling recruitment: \$70,000 Prize Pool for
Incentive: \$5,000 Data Processing and Cleaning: \$15,000 Miscellaneous
Expenses: \$5,000 Petty Cash: \$5,000

\hypertarget{idealized-questionnaire-and-survey}{%
\section{Idealized Questionnaire and
Survey}\label{idealized-questionnaire-and-survey}}

https://docs.google.com/forms/d/117wC8f1rBW\_bd5glEdiR-7823-02A3HiD5EzQuT1GVw/edit

\hypertarget{structure}{%
\section{Structure}\label{structure}}

\hypertarget{preamble-1}{%
\subsection{Preamble}\label{preamble-1}}

Title: 2024 Presidential Election Poll Data

Introduction: Thank you for your participation in our \$5,000 giveaway
program.

For entry, please complete the survey in its entirety below. Please be
informed that your answers are strictly confidential and will, under no
circumstances, be shared with any party, group, or persons of
affiliation.

\$5,000 Cash Voucher Program Details: Completing the survey below
qualifies as one entry into our giveaway. Each participant receives a
unique entry identifier, and every entry will have an equal chance in
our final random selection of the prize winner. Please note that each
individual may complete only one survey.

Referral Program: Our referral program offers you the chance to increase
your odds of winning. After completing the survey, you'll receive a
unique referral code. If a friend or associate completes the survey
using your code, your total number of entries doubles, effectively
doubling your chances of winning the \$5,000 prize.

Finally, we wish to clarify that this survey is conducted purely for
research purposes and is entirely unaffiliated with any political party,
organization, or individual in a position of power. Our goal is to gain
insight into the evolving political landscape and levels of public
engagement.

Referral Code (please ignore if inapplicable)

If you have entered a referral code above, please indicate the
relationship with the referee from the best choice below. Friend
Associate Relative Other\ldots{}

\hypertarget{eligibility-1}{%
\subsection{Eligibility}\label{eligibility-1}}

Do you meet your state's residency requirements? Yes No (End Survey)
Please select ``N/A'' from the options below if you reside in North
Dakota. Which of the following best describes you? I am registered to
vote I am not registered to vote I plan to register N/A Full Name (For
identification of \$5,000 cash winner) From which state will you be vote
for the next president of the United States of America? Where can we
reach you when we announce the winner? E-mail Telephone Other\ldots{}
Contact information **We will not use this for any other purpose than to
announce winners. It will only be used to reach our lucky winner for
instructions on receiving the cash prize. Which of the following age
groups best describes you? 18-25 26-31 32-40 41-50 51-60 61-80 81-125
Under 18 Which of the following best describes you? American Indian or
Alaskan Native Asian African American Hispanic or Latino Native Hawaiian
or Pacific Islander White Middle Eastern or North African Other\ldots{}
What is your gender? Male Female Prefer not to say Other\ldots{} Which
of the following situations best describe you? I did not vote in the
2020 presidential election I planned to vote but wasn't able to I voted
in the 2020 presidential election How likely are you to vote in the
upcoming election? Very unlikely 1 2 3 Very likely Have you attended a
presidential rally or their campaigns this year? Yes No Only in the past
On a scale of 1 to 3, how closely do you follow the presidential race?
Not at all 1 2 3 4 5 Very closely

\hypertarget{party-affiliation-1}{%
\subsection{Party Affiliation}\label{party-affiliation-1}}

Which party do you prefer a candidate to be chosen from? Democratic
Party's candidate Republican Party's candidate Another Party's candidate
In the 2020 presidential election, who did you vote for? If you did not
vote, who would you have voted for? Donald Trump Joe Biden Neither
Other\ldots{} If planning to vote, which presidential candidate are you
planning to vote for? If you are abstaining, please select which
candidate you feel more strongly about. Donald Trump Kamala Harris
Other\ldots{}

\newpage

\hypertarget{additional-data-details}{%
\section{Additional data details}\label{additional-data-details}}

\newpage

\hypertarget{sec-model-details}{%
\section{Model details}\label{sec-model-details}}

\hypertarget{posterior-predictive-check}{%
\subsection{Posterior predictive
check}\label{posterior-predictive-check}}

In Figure~\ref{fig-ppcheckandposteriorvsprior-1} we implement a
posterior predictive check. This shows that the posterior predictive
density closely follows the observed data density. Discrepancies between
the two suggest that the model may be missing some aspect of the data
structure or variability, which could guide potential model adjustments
or refinement.

In Figure~\ref{fig-ppcheckandposteriorvsprior-2} we compare the
posterior with the prior. This shows the prior for each parameter is
centered around zero, but the posterior shows more precise estimates for
each candidate and the date variable, reflecting the data's influence.
Also, the posterior narrowing suggests that the model successfully
extracted information from the data, reducing uncertainty around
parameter estimates compared to the prior. model\_data \textless-
read\_parquet(here::here(``data/02-analysis\_data/model\_data.parquet''))

\begin{figure}

\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{paper_files/figure-pdf/fig-ppcheckandposteriorvsprior-1.pdf}

}

}

\subcaption{\label{fig-ppcheckandposteriorvsprior-1}Posterior prediction
check}
\end{minipage}%
%
\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{paper_files/figure-pdf/fig-ppcheckandposteriorvsprior-2.pdf}

}

}

\subcaption{\label{fig-ppcheckandposteriorvsprior-2}Comparing the
posterior with the prior}
\end{minipage}%
\newline
\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{paper_files/figure-pdf/fig-ppcheckandposteriorvsprior-3.pdf}

}

}

\subcaption{\label{fig-ppcheckandposteriorvsprior-3}Posterior prediction
check}
\end{minipage}%

\caption{\label{fig-ppcheckandposteriorvsprior}Examining how the model
fits, and is affected by, the data}

\end{figure}

\hypertarget{diagnostics}{%
\subsection{Diagnostics}\label{diagnostics}}

Figure~\ref{fig-stanareyouokay-1} is a trace plot. It shows chain
mixing, stationarity and no drifting. This suggests that the model's
MCMC sampling achieved convergence for each parameter. The chains are
well-mixed and stationary, providing confidence in the posterior
estimates derived from this sampling process.

Figure~\ref{fig-stanareyouokay-2} is a Rhat plot. It shows the Rhat
values for the parameters estimated in the model. Values close to 1
indicate that the chains have mixed well and the posterior distribution
has been adequately explored. This suggests that the model has converged
successfully and the estimates are reliable for inference.

\begin{figure}

\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{paper_files/figure-pdf/fig-stanareyouokay-1.pdf}

}

}

\subcaption{\label{fig-stanareyouokay-1}Trace plot}
\end{minipage}%
%
\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{paper_files/figure-pdf/fig-stanareyouokay-2.pdf}

}

}

\subcaption{\label{fig-stanareyouokay-2}Rhat plot}
\end{minipage}%

\caption{\label{fig-stanareyouokay}Checking the convergence of the MCMC
algorithm}

\end{figure}

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Best2024}{}}%
Best, Ryan, Aaron Bycoffe, et al. 2024. {``National: President: General
Election: 2024 Polls.''} \emph{FiveThirtyEight}.
\url{https://projects.fivethirtyeight.com/polls/president-general/2024/national/}.

\leavevmode\vadjust pre{\hypertarget{ref-boyce_swing_states}{}}%
Boyce, Lily, Lazaro Gamio, Eli Murray, and Alicia Parlapiano. 2024.
{``Tracking the Swing States for Harris and Trump.''} 2024.
\url{https://www.nytimes.com/interactive/2024/us/elections/presidential-election-swing-states.html}.

\leavevmode\vadjust pre{\hypertarget{ref-bayesplot}{}}%
Gabry, Jonah, and Tristan Mahr. 2024. {``Bayesplot: Plotting for
Bayesian Models.''} \url{https://mc-stan.org/bayesplot/}.

\leavevmode\vadjust pre{\hypertarget{ref-rstanarm}{}}%
Goodrich, Ben, Jonah Gabry, Imad Ali, and Sam Brilleman. 2024.
{``Rstanarm: {Bayesian} Applied Regression Modeling via {Stan}.''}
\url{https://mc-stan.org/rstanarm/}.

\leavevmode\vadjust pre{\hypertarget{ref-kimball_polling}{}}%
Kimball, Spencer, Camille Mumford, and Matt Taglia. 2024. {``How
Pollsters Have Adapted to Changing Technology and Voters Who Don't
Answer the Phone.''} 2024.
\url{https://theconversation.com/how-pollsters-have-adapted-to-changing-technology-and-voters-who-dont-answer-the-phone-240283}.

\leavevmode\vadjust pre{\hypertarget{ref-here}{}}%
Müller, Kirill. 2020. \emph{Here: A Simpler Way to Find Your Files}.
\url{https://CRAN.R-project.org/package=here}.

\leavevmode\vadjust pre{\hypertarget{ref-citeR}{}}%
R Core Team. 2022. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-arrow}{}}%
Richardson, Neal, Ian Cook, Nic Crane, Dewey Dunnington, Romain
François, Jonathan Keane, Dragoș Moldovan-Grünfeld, Jeroen Ooms, Jacob
Wujciak-Jens, and Apache Arrow. 2024. \emph{Arrow: Integration to
'Apache' 'Arrow'}. \url{https://CRAN.R-project.org/package=arrow}.

\leavevmode\vadjust pre{\hypertarget{ref-psychology_survey_response}{}}%
Roger Tourangeau, Lance J. Rips, and Kenneth Rasinski. 2000. {``The
Psychology of Survey Response.''} In \emph{The Psychology of Survey
Response}, 23--61. Cambridge University Press.
\url{https://doi.org/10.1017/CBO9780511819322.003}.

\leavevmode\vadjust pre{\hypertarget{ref-nyt_poll}{}}%
The New York Times. 2024. \emph{You Ask, We Answer: How the Times/Siena
Poll Is Conducted}. The New York Times.
\url{https://www.nytimes.com/article/times-siena-poll-methodology.html\#link-7be11f8b}.

\leavevmode\vadjust pre{\hypertarget{ref-tidyverse}{}}%
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019.
{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\leavevmode\vadjust pre{\hypertarget{ref-readr}{}}%
Wickham, Hadley, Jim Hester, and Jennifer Bryan. 2023. \emph{Readr: Read
Rectangular Text Data}. \url{https://CRAN.R-project.org/package=readr}.

\leavevmode\vadjust pre{\hypertarget{ref-scales}{}}%
Wickham, Hadley, Thomas Lin Pedersen, and Dana Seidel. 2023.
\emph{Scales: Scale Functions for Visualization}.
\url{https://CRAN.R-project.org/package=scales}.

\leavevmode\vadjust pre{\hypertarget{ref-knitr}{}}%
Xie, Yihui. 2024. \emph{Knitr: A General-Purpose Package for Dynamic
Report Generation in r}. \url{https://yihui.org/knitr/}.

\leavevmode\vadjust pre{\hypertarget{ref-Zahn2024}{}}%
Zahn, Max. 2024. {``What's at Stake for Elon Musk in the Presidential
Election?''} \emph{ABC News}, November.
\url{https://abcnews.go.com/Business/stake-elon-musk-presidential-election/story?id=115143069}.

\leavevmode\vadjust pre{\hypertarget{ref-kableExtra}{}}%
Zhu, Hao. 2024. \emph{kableExtra: Construct Complex Table with 'Kable'
and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.

\end{CSLReferences}

\end{document}